---
title: "Plasma"
author: Federico Zertuche
date: 
output: rmarkdown::github_document
---


```{r}

library(tidyverse)
library(lubridate)
library(survival)

```

```{r}

plasma_simple <- read_csv('./plasma_simple.csv')

```


## Hospital Stays:


- 49 patients had stays of 7 or less days.
- Stays lasted 12.4 days in average (sd 8.18).


```{r}

plasma_simple %>%
    ggplot() +
    geom_histogram(aes(x = days_hospital)) +
    geom_vline(xintercept = 7, linetype = 2, size = 2, alpha = 0.5) +
    labs(x = 'stay duration (days)')

plasma_simple %>%
    filter(days_hospital <= 7) %>%
    count()

plasma_simple %>%
    summarise(mean(days_hospital), sd(days_hospital))

plasma_simple %>%
    ggplot() +
    geom_boxplot(aes(x = vital_st, y = days_hospital, fill = vital_st), alpha = 0.5) +
    theme(legend.position = "None") +
    labs(x = '', y = 'stay duration (days)')

```

## Predict hospital stay duration from saturation or respiratory frequency?

No.

```{r}

plasma_simple %>%
    ggplot(aes(x = fr_basal, y = days_hospital, color = vital_st)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE)

plasma_simple %>%
    ggplot(aes(x = sat_basal, y = days_hospital, color = vital_st)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE)

```


## First models:

- Differences were computed by (last measurement - basal).
- All models controll for age, sex, death or alive, hospital, 
  number of days from: symptoms to hospitalization and symptoms to treatment.

- Treatment has a positive effect on saturation difference and 
  respiratory frequency difference.


```{r}

sat_mod <- lm(sat_diff ~ age + sex + group + vital_st + hospital + 
   nd_sint_tto + nd_sint_ing, data = plasma_simple) 

sat_mod %>%
    summary()

fr_mod <- lm(fr_diff ~ age + sex + group + vital_st + hospital + 
   nd_sint_tto + nd_sint_ing, data = plasma_simple)

fr_mod %>%
    summary()

```

## Distributions per Hospital

```{r}

plasma_simple %>%
    ggplot() +
    geom_density(aes(sat_diff, fill = hospital), position = 'dodge', alpha = 0.3) +
    labs(x = 'saturation difference')

plasma_simple %>%
    ggplot() +
    geom_density(aes(fr_diff, fill = hospital), position = 'dodge', alpha = 0.3) +
    labs(x = 'respiratory frequency difference')

plasma_simple %>%
    ggplot() +
    geom_density(aes(days_hospital, fill = hospital), position = 'dodge', alpha = 0.3) +
    labs(x = 'stay duration (days)')


```

## Treatment per Hospital


```{r}

plasma_simple %>%
    ggplot() +
    geom_density(aes(nd_sint_tto, fill = hospital), position = 'dodge', alpha = 0.3) +
    labs(x = 'days from symptoms to treatment')


plasma_simple %>%
    ggplot() +
    geom_density(aes(nd_sint_ing, fill = hospital), position = 'dodge', alpha = 0.3) +
    labs(x = 'days from symptoms to hospitalization')

```

## Missing data.


```{r}

plasma_simple %>%
    filter(is.na(sat_final)) %>%
    count()

# missing data:
plasma_simple %>%
    group_by(hospital) %>%
    mutate(n_hosp = n()) %>%
    filter(is.na(sat_final)) %>%
    group_by(hospital, group) %>%
    summarise(n_na = n(), n_hosp = max(n_hosp)) %>%
    mutate(percent = 100*(n_na / n_hosp)) %>%
    select(hospital, group, percent) %>%
    pivot_wider(names_from = group, values_from = percent)

```

## Hospitalization Days and Probability of Survival


```{r}

glm(days_hospital ~ age + sex + group + vital_st + hospital + 
    nd_sint_tto + nd_sint_ing, data=plasma_simple, 
    family = poisson(link='log')) %>%
    summary()


coxph(Surv(days_hospital, status) ~ age + sex + group + hospital + 
    nd_sint_tto + nd_sint_ing + hta + obesity + dm, 
    data=plasma_simple) %>%
    summary()

coxph(Surv(days_hospital, status) ~  group, 
    data=plasma_simple) %>%
    summary()

survfit(Surv(days_hospital, status) ~ group, data=plasma_simple) %>%
    broom::tidy() %>%
    ggplot(aes(time, estimate, color = strata)) +
    geom_line() +
    geom_ribbon(aes(ymin=conf.low, ymax=conf.high), alpha=0.2)

```

